---
title: "DQe-c Report"
output: 
  html_document: 
    highlight: tango
    toc: yes
---


```{r, echo=FALSE, include=FALSE}
require(data.table);require(dplyr);require(ggplot2);require(gridExtra);require(rmarkdown);require(knitr):require(plotly);require(DT);require(treemap); require(visNetwork)
source("keys.R");rm(pss)
 
```

##Load and Test Details

This DQec report is generated from testing completeness in `r CDM` data from `r org` on `r Sys.Date()`. 

### Table 1. List and Status of Common Data Model (CDM) -- here, `r CDM` -- Tables in this load
The table bolow provides a list of CDM tables provided (and not provided) in the data load.

The source data this table and the following graphics in this section are being generated from is `r paste("tablelist_",CDM,"_",org,"_",as.character(format(Sys.Date(),"%d-%m-%Y")),".csv", sep="")`
```{r, echo=FALSE, fig.align='center', fig.width= 10}
dato <- read.csv(paste("reports/load_details_",CDM,"_",org,"_",as.character(format(Sys.Date(),"%d-%m-%Y")),".csv", sep=""))
dato$TotalSizeKB <- ifelse(is.null(dato$TotalSizeKB) | is.na(dato$TotalSizeKB), 0, dato$TotalSizeKB)
dato$TotalSizeGB <- round(as.numeric(dato$TotalSizeKB)/1000000,4)
dato$Rows <- ifelse(is.null(dato$Rows) | is.na(dato$Rows), 0, dato$Rows)
dato$status <- ifelse((dato$loaded == "Yes" & dato$available == "No"), "Table loaded but empty", "Table available")
dato$status <- ifelse((dato$loaded == "No"), "Table not loaded!", dato$status)
datatable(select(dato, CDM_Tables, status, TotalSizeGB, Rows), options = list(pageLength = 5), filter = 'bottom')

```


### Figure 1. Available Tables, Compared to all CDM (`r CDM`) Tables
This figure shows which of the CDM tables are loaded and/or available. 
```{r, echo=FALSE, fig.align='center', fig.width= 10}
treemap(dato,
        index=c("CDM_Tables"),
        vSize="index",
        vColor="status",
        type="categorical",
        title = "",
        title.legend = "Table Availability",
        border.col = "white",
        position.legend = "bottom"
        )

```

### Figure 2. File Size and Row Numbers by Table in the (`r CDM`) Load
```{r, echo=FALSE, fig.align='center', fig.width= 10}
treemap(dato,
        index=c("CDM_Tables"),
        vSize="Rows",
        vColor="TotalSizeGB",
        type="value",
        title = "",
        title.legend = "Size represents number of rows and color represent file size (in GB) for each table.",
        border.col = "white",
        position.legend = "bottom"
)
```


### Figure 3. Loaded tables against CDM (`r CDM`) Relational Model.
The figure below shows a network visualization of the CDM data model, as well as highlighting the tables that are available in this load (legend is the same as in Figure 1).
```{r, echo=FALSE, fig.align='center', fig.width= 10}
dat.net <- dato
if (CDM == "PCORNET3") {
  nodes <- data.frame(id = dat.net$X, 
                    label = dat.net$CDM_Tables, 
                    group = c(dat.net$status),
                    shadow = c(FALSE))
edges <- data.frame(from = c(4,4,4,4,4,4,4,4,4,4,4,4,4,#demographics's
                             7,7,7,7,7,7,7,#encounter's
                             12 #prescribing's
                             ), 
                    to = c(2,3,1,5,6,7,8,15,10,11,12,13,14,#to demographic
                           15,1,13,5,14,10,12,# to encounter through encounterid, enc_type, and providerid
                           6 # to prescribing through prescribingid
                           ))

visNetwork(nodes, edges) %>%
  visGroups(groupname = "Table loaded but empty", color = "#EE9572", shape = "circle") %>%
  visGroups(groupname = "Table available", color = "#00C5CD", shape = "circle") %>%
  visGroups(groupname = "Table not loaded!", color = "gray", shape = "circle") %>%
  visOptions(highlightNearest = TRUE, nodesIdSelection = T) %>% 
  visEdges(arrows = 'from', scaling = list(min = .5, max = 1)) %>%
  visInteraction(navigationButtons = T, dragView = FALSE, zoomView = FALSE)

} else
  if (CDM == "OMOPV5") {
#     nodes <- data.frame(id = dat.net$X, 
#                     label = dat.net$CDM_Tables, 
#                     group = c(dat.net$status),
#                     shadow = F)
# edges <- data.frame(from = c(14,14,14,14,14,14,14,14,14,14,14,#person
#                              17,17,17,17,17,#provider
#                              1,1, #caresite
#                              8,8, #drug exposure
#                              18,18,18,18, #visit occurence
#                              4, #condition occurence
#                              16 #procedure occurence
#                              ), 
#                     to = c(9,11,13,18,9,8,4,16,10,2,5,#person's
#                            1,8,4,16,10,#provider's
#                            9,12,#care site's
#                            7,6,#drug exposure's
#                            8, 4, 16, 10, # visit occurence's
#                            3, #condition occurence's
#                            15 #procedure occurence's
#                            ))
# 
# visGroups(groupname = "Table loaded but empty", color = "#EE9572", shape = "circle") %>%
#   visGroups(groupname = "Table available", color = "#00C5CD", shape = "circle") %>%
#   visGroups(groupname = "Table not loaded!", color = "gray", shape = "circle") %>%
#   visOptions(highlightNearest = TRUE, nodesIdSelection = T) %>% 
#   visEdges(arrows = 'from', scaling = list(min = .5, max = 1)) %>%
#   visInteraction(navigationButtons = T, dragView = FALSE, zoomView = FALSE)
  }


```

## Completeness Results
### Table 2. The Master Completeness Results Table
The table below provides results of completeness test at the value/cell level.

* `TabNam` = `r CDM`  table name
* `ColNam` = Column name
* `DQLVL` = Level of importance for completeness test. (`X`: Extremely Important, `H`: Highly Important, `L`:Low Importance)
* `FRQ` = Frequency of rows
* `UNIQFRQ` = Frequency of unique values in each column
* `MS1_FRQ` = Frequency of cells with NULL/NA values or empty strings in each column
* `MS2_FRQ` = Frequency of cells with characters in each column that don't represent meaningful data -- including, '+', '-', '_','#', '$', '*', '\', '?', '.', '&', '^', '%', '!', '@', and 'NI'.
* `MSs_PERC` = Percentage of overall missing data in each column 

Data for this table is generated from `r paste("DQ_Master_Table_",CDM,"_",org,"_",as.character(format(Sys.Date(),"%d-%m-%Y")),".csv", sep="")` saved under report directory.
```{r, echo=FALSE}
DQTBL <- read.csv(paste("reports/mstabs/DQ_Master_Table_",CDM,"_",org,"_",as.character(format(Sys.Date(),"%d-%m-%Y")),".csv", sep=""))
datatable(DQTBL[,c("TabNam","ColNam","DQLVL","test_date","FRQ","UNIQFRQ","MS1_FRQ","MS2_FRQ","MSs_PERC" )], options = list(pageLength = 10), filter = 'bottom')
```


### Figure 4. Changes in Primary Keys Across Loads
Figure below profiles changes in primary keys across loads as a measure of change in patient/record number over time.

Data for the figure is stored in `r paste("FRQ_comp_",usrnm,"_",as.character(format(Sys.Date(),"%d-%m-%Y")),".csv", sep="")`
```{r, echo=FALSE, warning=FALSE,  message=FALSE, fig.align='center', fig.height= 10,  fig.width= 10}
comp <- read.csv(paste("reports/FRQ_comp_",CDM,"_",org,"_",as.character(format(Sys.Date(),"%d-%m-%Y")),".csv", sep=""))
  
ggplot(comp, aes(x=reorder(test_date,test_date), y=UNIQFRQ, group = ColNam) ) + 
  geom_line(aes(), alpha = 0.4,  size = 1, show.legend = FALSE) + 
  stat_smooth(colour = "red",level=0.99) +
  geom_point( alpha = 0.5, shape = 1, size = 6, colour = "#FF3333", stroke = 2, show.legend = FALSE) + 
  geom_point( aes(col = test_date), shape = 20, size = 8,   show.legend = T) + 
  theme(plot.title = element_text(family = "Trebuchet MS", color="#666666", face="bold", hjust=0)) +
  xlab("Load") + ylab("Unique Counts") + 
  facet_wrap(~TabNam, ncol = 3, switch = "x", scales = "free") +
  theme(axis.text.x=element_text(colour="white", size = 0.1))
```

### Figure set 1. Proportion of Missing Data by Type in Loaded Tables
Figures below show proportion of missing cells/values in each column of each table loaded. Figures are generated based on Table 2. 

* `MS1_FRQ` = Frequency of cells with NULL/NA values and empty strings in each column -- presence of absence
* `MS2_FRQ` = Frequency of cells with characters in each column that don't represent meaningful data -- presence of nonsense 

```{r, echo=FALSE, ggplot, warning=FALSE,  message=FALSE, fig.align='center', fig.height= 5, fig.width= 12}

DQTBL$MS1_PERC <- ifelse(DQTBL$FRQ == 0, 0, round((DQTBL$MS1_FRQ)/DQTBL$FRQ,2))
DQTBL$MS2_PERC <- ifelse(DQTBL$FRQ == 0, 0, round((DQTBL$MS2_FRQ)/DQTBL$FRQ,2))
DF <- subset(melt(DQTBL, id.var=c("TabNam","ColNam")), variable %in% c("MS1_PERC","MS2_PERC"))
DF$value <- as.numeric(DF$value)
colnames(DF)[3] <- "DQ_Issue"
colnames(DF)[4] <- "Ratio"

#####plotting
plot_list<- list()
table_list <- unique(DQTBL$TabNam)

# create for loop to produce ggplot2 graphs 
for (i in seq_along(table_list)) { 
  
  # create plot for each OMOP table in DQTBL 
  plot_list[[i]] <- 
    plot_list[[i]] <- 
    ggplot(subset(DF, TabNam==table_list[i]),
           aes(x=ColNam, y=Ratio, fill=DQ_Issue)) + 
    geom_bar(stat="identity", width = 1) +
    #     scale_fill_manual(values=c("red","green","orange","dark red")) +
    facet_wrap( ~ ColNam, scale="free_x", nrow = 1) +
    #     ggtitle("Frequency of Missing Data") +
    xlab("Column") +
    ylab("Frequency") +
    theme(plot.title = element_text(family = "Trebuchet MS", color="#666666", face="bold", hjust=0),
          axis.text.x = element_text(vjust = 1),
          strip.text.x = element_text(angle = 90, face="bold")) +
    theme(axis.text.x=element_blank())+    # theme(legend.position="none") + 
    ggtitle(paste('Ratio of Missing Data in "', table_list[i], '" table', sep=''))
  
}

plot_list

```

##Data Model Tests

### Figure set 2. Common Key Variables
Figures below visualize number of unique key variables that are common in multiple `r CDM` tables. 

* The Reference column on the right comes from the table in which the variable is a primary key, and therefore is a reference for all other tables. 

* Count_Out shows number of unique key variables that are not present in the reference table -- e.g., person id from observation table that does not exist in person table.

* Count_In represent number of unique key variables that are present in the reference table -- e.g., person id from observation table that exist in person table as well.

```{r, echo=FALSE,  fig.align='center', fig.height= 3, fig.width= 10}
DQTBL_KEYS <- read.csv(paste("reports/DM_",CDM,"_",org,"_",as.character(format(Sys.Date(),"%d-%m-%Y")),".csv", sep=""))

if (CDM == "PCORNET3") {
  patid<- filter(DQTBL_KEYS, ColNam == "patid")
plot2 <- ggplot(data=patid, aes(x=reorder(TabNam, UNIQFRQ), y=UNIQFRQ)) +
  geom_bar(stat="identity", width = 1, aes(fill=Index)) +
  scale_fill_manual(values=c("darkolivegreen3","firebrick1", "springgreen3")) +
  ggtitle(paste("Count of Unique ",unique(patid$ColNam)," in Tables with ",unique(patid$ColNam),
                sep="")) +
  xlab(" Table Name") +
  ylab("Frequency of Unique Values")
(gg <- ggplotly(plot2))
} else
  if (CDM == "OMOPV5") {
#     person_id<- filter(DQTBL_KEYS, ColNam == "person_id")
# plot2 <- ggplot(data=person_id, aes(x=reorder(TabNam, UNIQFRQ), y=UNIQFRQ)) +
#   geom_bar(stat="identity", width = 1, aes(fill=Index)) +
#   scale_fill_manual(values=c("darkolivegreen3","firebrick1", "springgreen3")) +
#   ggtitle(paste("Count of Unique ",unique(person_id$ColNam)," in Tables with ",unique(person_id$ColNam),
#                 sep="")) +
#   xlab(" Table Name") +
#   ylab("Frequency of Unique Values")
# (gg <- ggplotly(plot2))
  }



```

```{r, echo=FALSE, fig.align='center', fig.height= 3, fig.width= 10}


if (CDM == "PCORNET3") {
  encounterid <- filter(DQTBL_KEYS, ColNam == "encounterid")

plot3 <- ggplot(data=encounterid, aes(x=reorder(TabNam, UNIQFRQ), y=UNIQFRQ)) +
  geom_bar(stat="identity", width = 1, aes(fill=Index)) +
  scale_fill_manual(values=c("darkolivegreen3","firebrick1", "springgreen3")) +
  ggtitle(paste("Count of Unique ",unique(encounterid$ColNam)," in Tables with ",unique(encounterid$ColNam),
                sep="")) +
  xlab(" Table Name") +
  ylab("Frequency of Unique Values")
(gg <- ggplotly(plot3))
} else
  if (CDM == "OMOPV5") {
#     care_site_id<- filter(DQTBL_KEYS, ColNam == "care_site_id")
# 
# plot3 <- ggplot(data=care_site_id, aes(x=reorder(TabNam, UNIQFRQ), y=UNIQFRQ)) +
#   geom_bar(stat="identity", width = 1, aes(fill=Index)) +
#   scale_fill_manual(values=c("darkolivegreen3","firebrick1", "springgreen3")) +
#   ggtitle(paste("Count of Unique ",unique(care_site_id$ColNam)," in Tables with ",unique(care_site_id$ColNam),
#                 sep="")) +
#   xlab(" Table Name") +
#   ylab("Frequency of Unique Values")
# (gg <- ggplotly(plot3))
  }


```

```{r, echo=FALSE, fig.align='center', fig.height= 3, fig.width= 10}

if (CDM == "PCORNET3") {
  providerid <- filter(DQTBL_KEYS, ColNam == "providerid")

plot4 <- ggplot(data=providerid, aes(x=reorder(TabNam, UNIQFRQ), y=UNIQFRQ)) +
  geom_bar(stat="identity", width = 1, aes(fill=Index)) +
  scale_fill_manual(values=c("darkolivegreen3","firebrick1", "springgreen3")) +
  ggtitle(paste("Count of Unique ",unique(providerid$ColNam)," in Tables with ",unique(providerid$ColNam),
                sep="")) +
  xlab(" Table Name") +
  ylab("Frequency of Unique Values")
(gg <- ggplotly(plot4))
} else
  if (CDM == "OMOPV5") {
#     visit_occurrence_id<- filter(DQTBL_KEYS, ColNam == "visit_occurrence_id")
# 
# plot4 <- ggplot(data=visit_occurrence_id, aes(x=reorder(TabNam, UNIQFRQ), y=UNIQFRQ)) +
#   geom_bar(stat="identity", width = 1, aes(fill=Index)) +
#   scale_fill_manual(values=c("darkolivegreen3","firebrick1", "springgreen3")) +
#   ggtitle(paste("Count of Unique ",unique(visit_occurrence_id$ColNam)," in Tables with ",unique(visit_occurrence_id$ColNam),
#                 sep="")) +
#   xlab(" Table Name") +
#   ylab("Frequency of Unique Values")
# (gg <- ggplotly(plot4))
  }



```

```{r, echo=FALSE, fig.align='center', fig.height= 3, fig.width= 10}

if (CDM == "PCORNET3") {
  enc_type <- filter(DQTBL_KEYS, ColNam == "enc_type")

plot5 <- ggplot(data=enc_type, aes(x=reorder(TabNam, UNIQFRQ), y=UNIQFRQ)) +
  geom_bar(stat="identity", width = 1, aes(fill=Index)) +
  scale_fill_manual(values=c("darkolivegreen3","firebrick1", "springgreen3")) +
  ggtitle(paste("Count of Unique ",unique(enc_type$ColNam)," in Tables with ",unique(enc_type$ColNam),
                sep="")) +
  xlab(" Table Name") +
  ylab("Frequency of Unique Values")
(gg <- ggplotly(plot5))
} else
  if (CDM == "OMOPV5") {
#     location_id<- filter(DQTBL_KEYS, ColNam == "location_id")
# 
# plot5 <- ggplot(data=location_id, aes(x=reorder(TabNam, UNIQFRQ), y=UNIQFRQ)) +
#   geom_bar(stat="identity", width = 1, aes(fill=Index)) +
#   scale_fill_manual(values=c("darkolivegreen3","firebrick1", "springgreen3")) +
#   ggtitle(paste("Count of Unique ",unique(location_id$ColNam)," in Tables with ",unique(location_id$ColNam),
#                 sep="")) +
#   xlab(" Table Name") +
#   ylab("Frequency of Unique Values")
# (gg <- ggplotly(plot5))
  }




```

```{r, echo=FALSE, fig.align='center', fig.height= 3, fig.width= 10}
if (CDM == "PCORNET3") {
  prescribingid <- filter(DQTBL_KEYS, ColNam == "prescribingid")

plot6 <- ggplot(data=prescribingid, aes(x=reorder(TabNam, UNIQFRQ), y=UNIQFRQ)) +
  geom_bar(stat="identity", width = 1, aes(fill=Index)) +
  scale_fill_manual(values=c("darkolivegreen3","firebrick1", "springgreen3")) +
  ggtitle(paste("Count of Unique ",unique(prescribingid$ColNam)," in Tables with ",unique(prescribingid$ColNam),
                sep="")) +
  xlab(" Table Name") +
  ylab("Frequency of Unique Values")
(gg <- ggplotly(plot6))
} else
  if (CDM == "OMOPV5") {
#     organization_id<- filter(DQTBL_KEYS, ColNam == "organization_id")
# 
# plot6 <- ggplot(data=organization_id, aes(x=reorder(TabNam, UNIQFRQ), y=UNIQFRQ)) +
#   geom_bar(stat="identity", width = 1, aes(fill=Index)) +
#   scale_fill_manual(values=c("darkolivegreen3","firebrick1", "springgreen3")) +
#   ggtitle(paste("Count of Unique ",unique(organization_id$ColNam)," in Tables with ",unique(organization_id$ColNam),
#                 sep="")) +
#   xlab(" Table Name") +
#   ylab("Frequency of Unique Values")
# (gg <- ggplotly(plot6))
  }



```



##Test of Completeness in Key Clinical Indicators

### Figure 5. Common Key Variables
Figure 5 shows the parcentage of patients missing specific key clinical indicators. 

```{r, echo=FALSE, fig.align='center', fig.height= 5, fig.width= 10}
withouts <- read.csv(paste("reports/withouts_",CDM,"_",org,"_",as.character(format(Sys.Date(),"%d-%m-%Y")),".csv", sep=""))
  
ggplot(withouts, aes(x=missing.percentage,y=reorder(group,-missing.percentage), label = perc)) + 
    geom_point(aes(fill = missing.percentage),shape = 21, colour = "black",  size = 8, stroke = 3, alpha = 0.9) +
    geom_text(vjust = -0.5, hjust = -.38, nudge_y = 0, size = 4)+
    scale_fill_gradient(limits=c(0, 100),low="#FFFAF0", high="#EE2C2C", guide=F,na.value="white")+
    labs(x = "", y = "") +
    scale_x_continuous(limits = c(0, 100)) + 
    theme_minimal() +
    theme(panel.grid.major.y = element_line(color = "gray",size = 2.5),
          panel.grid.major.x = element_line(colour = "black", linetype = "dotted"),
          axis.text.y=element_text(size=20, face="bold")) 
```


##info
This is report is from DQe-c version 3.1

For questions and/or inquiries email: `hestiri@mgh.harvard.edu`

